---
title: "WUP_04"
author: "Dorothy Decker!"
date: "Summer 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
thedata<-read.csv(file="GSScsv.csv")
```

```{r eval=FALSE}
thedata <- thedata %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

First, by looking at the data, we had to find out what sort of data we have. We used a past assignment, and pasted the data in and changed some stuff up, and found out we had numerical data. Which was WORDSUM, we now have a factor variable called DEGREE. We decided we have a problem called numerical~categorical type of problem. 

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
ggplot(thedata, aes(x = DEGREE, y = WORDSUM)) +
geom_boxplot(fill = "skyblue", color = "darkblue", outlier.shape = NA) +
theme_minimal() +
labs(x = "DEGREE", y = "WORDSUM") +
theme(plot.title = element_text(hjust = 0.5)) +
ggtitle("Fancy Box Plot")
```


According to the graph, we can visually understand there is a difference between the WORDSUM based on degree. A few of them look different from each other, and a few look like they may bump into each other. It looks like LT high school is different from graduate and junior college, and don't look like they would bump into each other. While Bachelor and high school look like they actually might bump into each other. This is all just by looking at the plot.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=thedata)
```

Using the numerical results of the average score for each group which is separated by degree, we can see there is a difference of scores. For example the group that scored the highest is the graduate. While the group that scored the lowest is lt high school. These two groups have the biggest difference of average statistically. Looking at these results, we can further understand that there may be difference of scores based on degrees.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=thedata)
summary.aov(model1)
```

```{r}
# Perform one-way ANOVA
model <- aov(WORDSUM ~ DEGREE, data = thedata)

# Conduct post-hoc test (Tukey's HSD)
posthoc <- TukeyHSD(model)

# Print the results
print(posthoc)
```


There is a significant difference between these categories. I reject the null hypothesis, based on degree, there is a difference on how people do on the test depending on their level of degree. What are the chances they are not different from each other? The probability is 0.000236, which is a very small number. Those chances are very slim, which can further lead us to conclude they are different. In the turkey test, the high school graduate has a p adj of 0.0037410 and LT high school graduate has a p adj of 0.0068774. These numbers show these two groups have the biggest difference of scores.



# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhaps are related for some other reason.  Watch the chapter 6 videos from the George Town videos collection.!-->
